---
title: "S&P Hypothesis Testing"
author: "Jon Kashmeeri"
date: "3/24/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Research Question

COVID 19 is affecting different sectors of the economy differently, in turn, causing many of the stock indexes on the market to fluctuate. For this research project, I  will be specifically testing the S&P 500 index compared to some of its subsections, in order see if their relationships are relevant. (the subsections are: communications, technologies, finance, and health care)


## Load Data

```{r loadData}

library(readr)

SP_500_dataset <- read_csv("^GSPC.csv")

SP_Comm_dataset <- read_csv("XLC.csv")

SP_Tech_dataset <- read_csv("XLK.csv")

SP_Fnan_dataset <- read_csv("XLF.csv")

SP_Health_dataset <- read_csv("XLV.csv")

```


## Load Packages

```{r loadPackages}

library('dplyr')
library('tidyverse')
library('TTR')

library(ggpubr)
library(rstatix)
```


##Wrangle Data

```{r}
SP_500_dataset <- SP_500_dataset %>%
  mutate(Date2 = Date)


fixed_500_dataset <- SP_500_dataset %>%
  separate (Date2, c("month","day","year"))

fixed_500_dataset <- fixed_500_dataset %>%
  unite("Date_id", `year`, `month`, `day`, sep = "-")

```

```{r}
SP_Comm_dataset <- SP_Comm_dataset %>%
  mutate(Date2 = Date)


fixed_Comm_dataset <- SP_Comm_dataset %>%
  separate (Date2, c("month","day","year"))

fixed_Comm_dataset <- fixed_Comm_dataset %>%
  unite("Date_id", `year`, `month`, `day`, sep = "-")

```

```{r}

SP_Tech_dataset <- SP_Tech_dataset %>%
  mutate(Date2 = Date)


fixed_Tech_dataset <- SP_Tech_dataset %>%
  separate (Date2, c("month","day","year"))

fixed_Tech_dataset <- fixed_Tech_dataset %>%
  unite("Date_id", `year`, `month`, `day`, sep = "-")

```

```{r}
SP_Fnan_dataset <- SP_Fnan_dataset %>%
  mutate(Date2 = Date)


fixed_Fnan_dataset <- SP_Fnan_dataset %>%
  separate (Date2, c("month","day","year"))

fixed_Fnan_dataset <- fixed_Fnan_dataset %>%
  unite("Date_id", `year`, `month`, `day`, sep = "-")

```

```{r}
SP_Health_dataset <- SP_Health_dataset %>%
  mutate(Date2 = Date)


fixed_Health_dataset <- SP_Health_dataset %>%
  separate (Date2, c("month","day","year"))

fixed_Health_dataset <- fixed_Health_dataset %>%
  unite("Date_id", `year`, `month`, `day`, sep = "-")

```


## Normalizing data

```{r}
normalized_500_dataset <- fixed_500_dataset %>%
  select(Date, `Open`, `Adj Close`, `High`, `Low`, `Volume`,`Date_id`) 

normalized_500_dataset$NormalOpen <- ((normalized_500_dataset$Open/normalized_500_dataset$Open[1])*100)
normalized_500_dataset$NormalClose <- ((normalized_500_dataset$`Adj Close`/normalized_500_dataset$`Adj Close`[1])*100)
normalized_500_dataset$NormalHigh <- ((normalized_500_dataset$High/normalized_500_dataset$High[1])*100)
normalized_500_dataset$NormalLow <- ((normalized_500_dataset$Low/normalized_500_dataset$Low[1])*100)
normalized_500_dataset$NormalVolume <- ((normalized_500_dataset$Volume/normalized_500_dataset$Volume[1])*100)

normalized_500_dataset <- normalized_500_dataset %>%
  select(7:12)

```

```{r}
normalized_Comm_dataset <- fixed_Comm_dataset %>%
  select(Date, `Open`, `Adj Close`, `High`, `Low`, `Volume`, `Date_id`) 

normalized_Comm_dataset$NormalOpen <- ((normalized_Comm_dataset$Open/normalized_Comm_dataset$Open[1])*100)
normalized_Comm_dataset$NormalClose <- ((normalized_Comm_dataset$`Adj Close`/normalized_Comm_dataset$`Adj Close`[1])*100)
normalized_Comm_dataset$NormalHigh <- ((normalized_Comm_dataset$High/normalized_Comm_dataset$High[1])*100)
normalized_Comm_dataset$NormalLow <- ((normalized_Comm_dataset$Low/normalized_Comm_dataset$Low[1])*100)
normalized_Comm_dataset$NormalVolume <- ((normalized_Comm_dataset$Volume/normalized_Comm_dataset$Volume[1])*100)

normalized_Comm_dataset <- normalized_Comm_dataset %>%
  select(7:12)

```

```{r}
normalized_Tech_dataset <- fixed_Tech_dataset %>%
  select(Date, `Open`, `Adj Close`, `High`, `Low`, `Volume`, `Date_id`) 

normalized_Tech_dataset$NormalOpen <- ((normalized_Tech_dataset$Open/normalized_Tech_dataset$Open[1])*100)
normalized_Tech_dataset$NormalClose <- ((normalized_Tech_dataset$`Adj Close`/normalized_Tech_dataset$`Adj Close`[1])*100)
normalized_Tech_dataset$NormalHigh <- ((normalized_Tech_dataset$High/normalized_Tech_dataset$High[1])*100)
normalized_Tech_dataset$NormalLow <- ((normalized_Tech_dataset$Low/normalized_Tech_dataset$Low[1])*100)
normalized_Tech_dataset$NormalVolume <- ((normalized_Tech_dataset$Volume/normalized_Tech_dataset$Volume[1])*100)

normalized_Tech_dataset <- normalized_Tech_dataset %>%
  select(7:12)

```

```{r}
normalized_Fnan_dataset <- fixed_Fnan_dataset %>%
  select(Date, `Open`, `Adj Close`, `High`, `Low`, `Volume`, `Date_id`) 

normalized_Fnan_dataset$NormalOpen <- ((normalized_Fnan_dataset$Open/normalized_Fnan_dataset$Open[1])*100)
normalized_Fnan_dataset$NormalClose <- ((normalized_Fnan_dataset$`Adj Close`/normalized_Fnan_dataset$`Adj Close`[1])*100)
normalized_Fnan_dataset$NormalHigh <- ((normalized_Fnan_dataset$High/normalized_Fnan_dataset$High[1])*100)
normalized_Fnan_dataset$NormalLow <- ((normalized_Fnan_dataset$Low/normalized_Fnan_dataset$Low[1])*100)
normalized_Fnan_dataset$NormalVolume <- ((normalized_Fnan_dataset$Volume/normalized_Fnan_dataset$Volume[1])*100)

normalized_Fnan_dataset <- normalized_Fnan_dataset %>%
  select(7:12)

```

```{r}
normalized_Health_dataset <- fixed_Health_dataset %>%
  select(Date, `Open`, `Adj Close`, `High`, `Low`, `Volume`, `Date_id`) 

normalized_Health_dataset$NormalOpen <- ((normalized_Health_dataset$Open/normalized_Health_dataset$Open[1])*100)
normalized_Health_dataset$NormalClose <- ((normalized_Health_dataset$`Adj Close`/normalized_Health_dataset$`Adj Close`[1])*100)
normalized_Health_dataset$NormalHigh <- ((normalized_Health_dataset$High/normalized_Health_dataset$High[1])*100)
normalized_Health_dataset$NormalLow <- ((normalized_Health_dataset$Low/normalized_Health_dataset$Low[1])*100)
normalized_Health_dataset$NormalVolume <- ((normalized_Health_dataset$Volume/normalized_Health_dataset$Volume[1])*100)

normalized_Health_dataset <- normalized_Health_dataset %>%
  select(7:12)

```


## Normal Test Plots

```{r}
ggplot()+
  geom_point(aes(as.Date(Date_id),NormalOpen, color= "Normal 500"), shape = "circle", data=normalized_500_dataset)+
  geom_line(aes(as.Date(Date_id),NormalOpen), color= "Red", data=normalized_500_dataset)+
  
  geom_point(aes(as.Date(Date_id),NormalOpen, color= "Normal Comm"), shape = "triangle", data=normalized_Comm_dataset)+
  geom_line(aes(as.Date(Date_id),NormalOpen), color= "Blue", data=normalized_Comm_dataset)+ 
 
  geom_point(aes(as.Date(Date_id),NormalOpen, color= "Normal Tech"), shape = "square", data=normalized_Tech_dataset)+
  geom_line(aes(as.Date(Date_id),NormalOpen), color= "Green", data=normalized_Tech_dataset)+
  
  geom_point(aes(as.Date(Date_id),NormalOpen, color= "Normal Fnan"), shape = "plus", data=normalized_Fnan_dataset)+
  geom_line(aes(as.Date(Date_id),NormalOpen), color= "Black", data=normalized_Fnan_dataset)+
  
  geom_point(aes(as.Date(Date_id),NormalOpen, color= "Normal Health"), shape = "diamond", data=normalized_Health_dataset)+
  geom_line(aes(as.Date(Date_id),NormalOpen), color= "Yellow", data=normalized_Health_dataset)+
  
  labs(title = "All Daily NormalOpens", x = "Time", y = "NormalOpen", colors = "Legends")
  
```
  
  
## Moving Average Test

```{r}
normalized_500_dataset$MAOpen <- SMA(x = normalized_500_dataset$NormalOpen,n = 10) 

ggplot(normalized_500_dataset)+
  geom_line(mapping = aes(x = as.Date(Date_id), y = NormalOpen, color = "Normal S&P 500"))+
  geom_line(mapping = aes(x = as.Date(Date_id), y = MAOpen, color = "Moving Average"))+
  labs(title = "S&P 500 vs. Moving Average", x = "Time", y = "Open", colors = "Legends")

```

```{r}
normalized_Comm_dataset$MAOpen <- SMA(x = normalized_Comm_dataset$NormalOpen,n = 10) 

ggplot(normalized_Comm_dataset)+
  geom_line(mapping = aes(x = as.Date(Date_id), y = NormalOpen, color = "Normal S&P Comm"))+
  geom_line(mapping = aes(x = as.Date(Date_id), y = MAOpen, color = "Moving Average"))+
  labs(title = "S&P Comm vs. Moving Average", x = "Time", y = "Open", colors = "Legends")

```

```{r}
normalized_Tech_dataset$MAOpen <- SMA(x = normalized_Tech_dataset$NormalOpen,n = 10) 

ggplot(normalized_Tech_dataset)+
  geom_line(mapping = aes(x = as.Date(Date_id), y = NormalOpen, color = "Normal S&P Tech"))+
  geom_line(mapping = aes(x = as.Date(Date_id), y = MAOpen, color = "Moving Average"))+
  labs(title = "S&P Tech vs. Moving Average", x = "Time", y = "Open", colors = "Legends")

```

```{r}
normalized_Fnan_dataset$MAOpen <- SMA(x = normalized_Fnan_dataset$NormalOpen,n = 10) 

ggplot(normalized_Fnan_dataset)+
  geom_line(mapping = aes(x = as.Date(Date_id), y = NormalOpen, color = "Normal S&P Fnan"))+
  geom_line(mapping = aes(x = as.Date(Date_id), y = MAOpen, color = "Moving Average"))+
  labs(title = "S&P Fnan vs. Moving Average", x = "Time", y = "Open", colors = "Legends")

```

```{r}
normalized_Health_dataset$MAOpen <- SMA(x = normalized_Health_dataset$NormalOpen,n = 10) 

ggplot(normalized_Health_dataset)+
  geom_line(mapping = aes(x = as.Date(Date_id), y = NormalOpen, color = "Normal S&P Health"))+
  geom_line(mapping = aes(x = as.Date(Date_id), y = MAOpen, color = "Moving Average"))+
  labs(title = "S&P Health vs. Moving Average", x = "Time", y = "Open", colors = "Legends")

```


## T-Statistic 

```{r}
AllOpens <- normalized_500_dataset %>%
  select(NormalOpen) 
  AllOpens$SP_500_NormalOpen <- normalized_500_dataset$NormalOpen 
AllOpens <- AllOpens %>%
  select(-NormalOpen)

AllOpens[2] <- normalized_Comm_dataset %>%
  select(NormalOpen) 
  AllOpens$SP_Comm_NormalOpen <- normalized_Comm_dataset$NormalOpen 
AllOpens <- AllOpens %>%
  select(-NormalOpen)

AllOpens[3] <- normalized_Tech_dataset %>%
  select(NormalOpen) 
  AllOpens$SP_Tech_NormalOpen <- normalized_Tech_dataset$NormalOpen 
AllOpens <- AllOpens %>%
  select(-NormalOpen)

AllOpens[4] <- normalized_Fnan_dataset %>%
  select(NormalOpen) 
  AllOpens$SP_Fnan_NormalOpen <- normalized_Fnan_dataset$NormalOpen 
AllOpens <- AllOpens %>%
  select(-NormalOpen)

AllOpens[5] <- normalized_Health_dataset %>%
  select(NormalOpen) 
  AllOpens$SP_Health_NormalOpen <- normalized_Health_dataset$NormalOpen 
AllOpens <- AllOpens %>%
  select(-NormalOpen)



AllOpens[6] <- normalized_500_dataset %>%
  select(MAOpen) 
  AllOpens$normalized_500_MAOpen <- normalized_500_dataset$MAOpen 
AllOpens <- AllOpens %>%
  select(-MAOpen)

AllOpens[7] <- normalized_Comm_dataset %>%
  select(MAOpen) 
  AllOpens$normalized_Comm_MAOpen <- normalized_Comm_dataset$MAOpen 
AllOpens <- AllOpens %>%
  select(-MAOpen)

AllOpens[8] <- normalized_Tech_dataset %>%
  select(MAOpen) 
  AllOpens$normalized_Tech_MAOpen <-normalized_Tech_dataset$MAOpen 
AllOpens <- AllOpens %>%
  select(-MAOpen)

AllOpens[9] <- normalized_Fnan_dataset %>%
  select(MAOpen) 
  AllOpens$normalized_Fnan_MAOpen <-normalized_Fnan_dataset$MAOpen 
AllOpens <- AllOpens %>%
  select(-MAOpen)

AllOpens[10] <- normalized_Health_dataset %>%
  select(MAOpen) 
  AllOpens$normalized_Health_MAOpen <-normalized_Health_dataset$MAOpen 
AllOpens <- AllOpens %>%
  select(-MAOpen)

```

```{r}
m1<- mean(AllOpens$normalized_500_MAOpen, na.rm = TRUE)
m2 <- mean(AllOpens$normalized_Comm_MAOpen, na.rm = TRUE)

sd1<- sd(AllOpens$normalized_500_MAOpen, na.rm = TRUE)
sd2 <- sd(AllOpens$normalized_Comm_MAOpen, na.rm = TRUE)

```

```{r}
t.test(AllOpens$normalized_500_MAOpen, AllOpens$normalized_Health_MAOpen, na.rm = TRUE)

```




